---
title: "因果分析hw3"
author: "陳葳芃 統計碩1 314657018"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
source("renv/activate.R")
library(markdown)
```

## Part I: Estimation for continuous outcome. Show your code and answer by R Markdown.
### Q1-(A) According to the following models, please simulate a dataset with 1,000 observations by R, and show the first 5 observations.

```{r}
# 設定隨機種子以便重現
set.seed(123)
# 生成 L
L <- runif(1000, min = 8, max = 16)

# 定義 expit 函數
expit <- function(x) {
  1 / (1 + exp(-x))
}

# 計算 p_A
p_A <- expit(6 - 0.5 * L)
# 根據 Bernoulli 生成 A
A <- rbinom(1000, size = 1, prob = p_A)
# 計算 E[Y|A,L]
EY <- 20 + 2*A + 0.1*L
# 根據正態分布生成 Y
Y <- rnorm(1000, mean = EY, sd = 1)
# 建立資料框
df <- data.frame(L = L, A = A, Y = Y)
# 顯示前 5 筆資料
head(df, 5)
```

### Q1-(B)Suppose L represents children’s age, A represents whether a child walks to school (A = 0) or picked up by parents (A = 1), and  represents children’s BMI. Please derive the estimate of causal effect based on standardization method under difference scale (both point and 95% CI).

```{r}
lm.Y.AC = lm(Y ~ ., data = df)
newdata_df = df
newdata_df$A = 1
phi.1 = mean(predict(lm.Y.AC,newdata = newdata_df))
newdata_df$A = 0
phi.0 = mean(predict(lm.Y.AC,newdata = newdata_df))
print(phi.1 - phi.0)#point estimate by standardization method


B.lis = rep(NA, 3000)
for (b in 1:3000) {
  resample.df = df[sample(220,220,replace = T),]
  lm.Y.AC = lm(Y ~ ., data = resample.df)

  resample.df$A = 1
  phi.1 = mean(predict(lm.Y.AC,newdata = resample.df))

  resample.df$A = 0
  phi.0 = mean(predict(lm.Y.AC,newdata = resample.df))
  B.lis[b] = phi.1 - phi.0
  }
  quantile(B.lis,probs = c(0.025,0.975))#percentile bootstrap confidence interval of causal effect
```

### Q1-(C)Please derive the estimate of causal effect based on regression-based estimator under difference scale (both point and 95% CI).

```{r}
set.seed(123)
lm.Y.AC_Reg = lm(Y ~ ., data = df)
print(lm.Y.AC_Reg$coefficients[3])#point estimate by regression-based method


B_Reg.lis = rep(NA, 3000)
for (b in 1:3000) {
  resample.regdf = df[sample(220,220,replace = T),]
  lm.Y.AC_Reg = lm(Y ~ ., data = resample.regdf)

  B_Reg.lis[b] = lm.Y.AC_Reg$coefficients[3]
  }
quantile(B_Reg.lis,probs = c(0.025,0.975))#percentile bootstrap confidence interval of causal effect
```

### Q1-(D)Please derive the estimate of causal effect based on IPW estimator under difference scale (both point and 95% CI).

```{r}
lm.A.C = glm(A ~ .,data = df[,-2], family = binomial(link = "logit"))
phi.1 = mean((df$A == 1)*df$Y/predict(lm.A.C,type = "response"))
phi.0 = mean((df$A == 0)*df$Y/(1 - predict(lm.A.C,type = "response")))
print(phi.1 - phi.0)#IPW estimates of causal effect under difference scale


B.lis = rep(NA, 3000)
for (b in 1:3000) {
  resample.df = df[sample(220,220,replace = T),]
  lm.A.C = glm(A ~ .,data = resample.df[,1:2], family = binomial(link = "logit"))
  phi.1 = mean((resample.df$A == 1)*resample.df$Y/predict(lm.A.C,type = "response"))
  phi.0 = mean((resample.df$A == 0)*resample.df$Y/(1 - predict(lm.A.C,type = "response")))

  B.lis[b] = phi.1 - phi.0
  }
quantile(B.lis, probs = c(0.025,0.975))#percentile bootstrap confidence interval of causal effect


library(markdown)
knitr::spin("hw3.r", knit = TRUE)
```

